Universal Properties in Generalized Multicategories

### 2018-02-01 upgm.tex
most of what I have on left universal props, need 2-cells

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newcommand{\pto}{\nrightarrow}
\newcommand{\pfrom}{\nleftarrow}
\newcommand{\vcat}{\mathcal}
\newcommand{\cat}{\mathbb}
\newcommand{\vtkmnd}{\mathbb{K}\text{Mod}(\vcat{V},T)}
\newcommand{\rmod}{\text{RMod}}

@@

\begin{definition}{Profunctor of Right Modules}
  The profunctor of category-parameterized right modules on
  multicategories $\rmod : \vtkmnd \pfrom \vcat V$ has as elements
  $\vcat M \pfrom \cat C$ \emph{right modules} which consist of
  \begin{enumerate}
  \item A profunctor $Q : T \vcat M \pfrom \cat C$ in $\vcat V$.
  \item An action
    \[
\begin{tikzcd}
T^2M_0 \arrow[dd, "\mu"] & T M_0 \arrow[l, "T(M_1)"] & C \arrow[dd, "id"] \arrow[l, "Q"] \\
 & Qcut &  \\
TM_0 &  & C \arrow[ll, "Q"]
\end{tikzcd}
\]
  \item That is associative:
    \[
    \begin{tikzcd}
T^3M_0 \arrow[dd, "\mu"] &  & T^2M_0 \arrow[dd, "\mu"] \arrow[ll, "T^2(M_1)"] & T M_0 \arrow[l, "T(M_1)"] & C \arrow[dd, "id"] \arrow[l, "Q"] &  & T^3M_0 \arrow[dd, "\mu"] & T^2M_0 \arrow[l] & TM_0 \arrow[l] \arrow[dd, "id"] & C \arrow[l, "Q"] \arrow[dd, "id"] \\
 & \mu &  & Qcut &  &  &  & T(M_1) &  &  \\
T^2M_0 \arrow[dd, "\mu"] &  & TM_0 \arrow[ll, "T(M_1)"] &  & C \arrow[ll, "Q"] \arrow[dd, "id"] & = & T^2M_0 \arrow[dd, "\mu"] &  & TM_0 \arrow[ll, "T(M_1)"] & C \arrow[l, "Q"] \arrow[dd, "id"] \\
 &  & Qcut &  &  &  &  &  & Qcut &  \\
TM_0 &  &  &  & C \arrow[llll, "Q"] &  & TM_0 &  &  & C \arrow[lll, "Q"]
\end{tikzcd}
    \]
  \item And unital:
    \[\begin{tikzcd}
 & TM_0 \arrow[ld, "T\eta"] \arrow[d, "id"] &  & C \arrow[ll, "Q"] \arrow[d, "id"] &  &  &  \\
T^2M_0 \arrow[dd, "\mu"] & TM_0 \arrow[l, "M_1"] &  & C \arrow[ll, "Q"] \arrow[dd, "id"] & = & TM_0 \arrow[d, "id"] & C \arrow[d, "id"] \arrow[l, "Q"] \\
 &  & Qcut &  &  & TM_0 & C \arrow[l, "Q"] \\
TM_0 &  &  & C \arrow[lll, "Q"] &  &  & 
    \end{tikzcd}\]
  \end{enumerate}

@@

>>>>

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newcommand{\pto}{\nrightarrow}
\newcommand{\pfrom}{\nleftarrow}
\newcommand{\vcat}{\mathcal}
\newcommand{\cat}{\mathbb}
\newcommand{\vtkmnd}{\mathbb{K}\text{Mod}(\vcat{V},T)}
\newcommand{\rmod}{\text{RMod}}
\newcommand{\lmod}{\text{LMod}}

@@

\begin{definition}{Profunctor of Right Modules}
  The profunctor of category-parameterized right modules on
  multicategories $\rmod : \vtkmnd \pfrom \vcat V$ has as elements
  $\vcat M \pfrom \cat C$ \emph{right modules} which consist of
  \begin{enumerate}
  \item A profunctor $Q : T \vcat M_0 \pfrom \cat C$ in $\vcat V$.
  \item An action
    \[
\begin{tikzcd}
T^2M_0 \arrow[dd, "\mu"] & T M_0 \arrow[l, "T(M_1)"] & C \arrow[dd, "id"] \arrow[l, "Q"] \\
 & Qcut &  \\
TM_0 &  & C \arrow[ll, "Q"]
\end{tikzcd}
\]
  \item That is associative:
    \[
    \begin{tikzcd}
T^3M_0 \arrow[dd, "\mu"] &  & T^2M_0 \arrow[dd, "\mu"] \arrow[ll, "T^2(M_1)"] & T M_0 \arrow[l, "T(M_1)"] & C \arrow[dd, "id"] \arrow[l, "Q"] &  & T^3M_0 \arrow[dd, "\mu"] & T^2M_0 \arrow[l] & TM_0 \arrow[l] \arrow[dd, "id"] & C \arrow[l, "Q"] \arrow[dd, "id"] \\
 & \mu &  & Qcut &  &  &  & T(M_1) &  &  \\
T^2M_0 \arrow[dd, "\mu"] &  & TM_0 \arrow[ll, "T(M_1)"] &  & C \arrow[ll, "Q"] \arrow[dd, "id"] & = & T^2M_0 \arrow[dd, "\mu"] &  & TM_0 \arrow[ll, "T(M_1)"] & C \arrow[l, "Q"] \arrow[dd, "id"] \\
 &  & Qcut &  &  &  &  &  & Qcut &  \\
TM_0 &  &  &  & C \arrow[llll, "Q"] &  & TM_0 &  &  & C \arrow[lll, "Q"]
\end{tikzcd}
    \]
  \item And unital:
    \[\begin{tikzcd}
 & TM_0 \arrow[ld, "T\eta"] \arrow[d, "id"] &  & C \arrow[ll, "Q"] \arrow[d, "id"] &  &  &  \\
T^2M_0 \arrow[dd, "\mu"] & TM_0 \arrow[l, "M_1"] &  & C \arrow[ll, "Q"] \arrow[dd, "id"] & = & TM_0 \arrow[d, "id"] & C \arrow[d, "id"] \arrow[l, "Q"] \\
 &  & Qcut &  &  & TM_0 & C \arrow[l, "Q"] \\
TM_0 &  &  & C \arrow[lll, "Q"] &  &  & 
    \end{tikzcd}\]
  \end{enumerate}

@@

For left-universal properties, consider sums as an illustrative
example.
%
In a category $\cat C$, a sum/coproduct $A + B$ is defined by the property that
\[ \cat C(A + B, C) \equiv \cat C(A,C) \times \cat C(B,C) \]
%
However, a sum in a multicategory $\vcat C$ should \emph{not} just be
a sum in the underlying category (unlike for right universal
properties).
%
Instead it is what's called a \emph{distributive} coproduct, which can
be defined manually by some laws relating compositions of the
coproduct with the monoidal structure.
%
We want to find an abstract setting that captures these distributivity
properties, which already have a type-theoretic explanation: the
equivalence happens ``in an arbitrary context'':
\[ \vcat C(\Gamma,A + B; C) \equiv \vcat C(\Gamma, A; C) \times C(\Gamma,B;C)\]
this can be defined using fibrations but we will try a different,
hopefully more direct tactic.

We define left universal properties as saying that maps out of a
context with the object are given by a certain module. For this to
make sense we need a concept of ``context with an object missing'',
and after McBride and Combinatorialists we will call this the
``derivative'' of the context. Since the shape of contexts is given by
a monad, we would like a concept that applies to an arbitrary monad,
not just polynomial functors. I haven't nailed it down but it looks
like we need the following at least.

The microcoms principle is here inescapable and so we need an extra
condition of $\vcat V$ being a \emph{monoidal} virtual double
category, which we will here weaken to a \emph{virtually monoidal}
double category which might be called a ``virtual multi-double
category'' or something like that.

\begin{definition}[Derivative of a Monad]
  Let $T$ be a monad on a virtually monoidal virtual equipment $\vcat
  V$. A \emph{derivative} of $T$ is an endofunctor $dT : \vcat V \to
  \vcat V$ such that/with
  \begin{enumerate}
  \item $dT$ is a vertical monoid in $\vcat V$, i.e. there are natural
    monoidal product $m_A : dT(A),dT(A) \to dT(A)$ and unit $e_A :
    \cdot \to dT(A)$.
  \item $dT$ is a left-module of $T$, i.e. there is a vertical action
    $d\mu : dT \circ T \Rightarrow dT$ that interacts with $\mu,\eta$
    in the obvious way.
  \item A ``plugging in'' morphism $dT,Id \Rightarrow T$, i.e., $l :
    dT(A),A \to TA$.
  \item $\lambda l : dT \to (Id \to T)$ is a monoid homomorphism
    where $(Id \to T)$ uses kleisli composition
  \item The two ways of plugging $dT(T A), A \to T A$ are equal:
    \[\begin{tikzcd}
 & dT(TA),A \arrow[ld, "{dT(\eta),id}"'] \arrow[rd, "{id,\eta}"] &  \\
dT(A),A \arrow[rdd, "l_A"] &  & dT(TA),TA \arrow[d, "l_{TA}"] \\
 &  & T^2A \arrow[ld, "\mu"] \\
 & TA & 
    \end{tikzcd}\]
  \item Possibly other laws...
  \end{enumerate}
\end{definition}
What other conditions we have no clue unfortunately. For instance all
of the above are satisfied by the trivial functor $dT(A) = 1$

Given a derivative $dT$ of $T$, we define the profunctor of
\emph{left} modules as follows:

\begin{definition}[Profunctor of Left Modules]
  The profunctor of category-parameterized left modules on
  multicategories $\lmod : \vcat V \pfrom \vtkmnd$ has as elements
  $\cat C \pfrom \vcat M$ \emph{left modules} which consist of
  \begin{enumerate}
  \item A profunctor $P : dT(\vcat M_0) \otimes \cat C \pfrom \vcat M_0$ in $\vcat V$
  \item An action \[\begin{tikzcd}
dT(TM_0) \times (dT(M_0) \times C) \arrow[dd, "l"] &  & dT(M_0) \times C \arrow[ll, "dT(M_1) \times P"] &  & M_0 \arrow[dd, "id"] \arrow[ll, "P"] \\
 &  & Pcut &  &  \\
dT(M_0) \times C &  &  &  & M_0 \arrow[llll, "P"]
  \end{tikzcd}\]
  \item That is associative
  \item Unital
  \item 2-cells are defined as...
  \end{enumerate}
\end{definition}
